Binning Strategies and Related Loss for Large Data

### 2015-12-15 versionForJCGS/PaperJCGS.Rnw
Reorganize data structure in folder

# connect <- dbConnect(dbDriver("MySQL"), host="mysql2.stat.iastate.edu", 
#                      port=3306, user="dbaccess", dbname="baseball")
# pitch <- new("dataDB", co=connect, table="Pitching")
# 
# d1 <- dbData(pitch, vars=c( "G", "SO")) # Games, Strike Outs
# 
# xtd.data <- function(x) {
# # replicate each line as often as given by the last entry
#   k <- dim(x)[2]
# 
#   res <- data.frame(sapply(1:(k-1), function(var) {
#   		return(rep(x[,var], x[,k]))
# 		}
# 	))
# 	names(res) <- names(x)[1:(k-1)]
# 	return(res)
# }
# d1.ext <- xtd.data(d1)
# dbDisconnect(connect)
# pitchdat <- d1.ext
# save(d1, d1.ext,pitchdat, file="pitching.Rdata")
load("pitching.Rdata")
@

@@

<<OriginOffsetSimulations, echo=F,include=T,eval=T,fig.width=8, fig.height=4, out.width='.9\\linewidth', fig.pos='h',fig.align='center',tidy=F, cache=TRUE, fig.show='hold',fig.cap="Net spatial loss for coarse simulated data at a 2X2 unit resolution using various sized square bins over the range of possible origin offsets. The vertical gray line in each facet indicates the origin offset of (1,1).">>= 
##  Check out the spatial loss for the data that is recorded to resolution = alpha at various origin offsets
# standard and random binning with bin sizes that are scalar multiples of the data resolution
# save spatial loss information
# spatlossdatRound <- data.frame(NULL)
# index=0
# distindex = 0
# distrib <- c("Uniform", "Normal", "Exponential")
# widths <- seq(alpha,5*alpha,by=alpha)
# for (dat in 1:3){
#   distindex <- distindex + 1
#   for (width in widths){
#     for (origin in seq(-width,0, by=width/60)){
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="standard")
#       binout[[2]]$bintype <- "standard"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="random")
#       binout[[2]]$bintype <- "random"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       print(c(distrib[distindex],width,origin))
#     }
#   }
# }
# write.csv(spatlossdatRound,"LossesByOriginBinwidth.csv", row.names=FALSE)
lossOrigin <- read.csv("LossesByOriginBinwidth.csv",header=TRUE)
mins <- ddply(lossOrigin, .(widthx, bintype, distrib), summarise,
              minx = max(originx[which(totalSpatialLoss == min(totalSpatialLoss))]),
              minloss = min(totalSpatialLoss))
lossOrigin$binsize <- factor(paste(lossOrigin$widthx,"X",lossOrigin$widthx,sep=""),
                             levels=c("2X2","4X4","6X6","8X8","10X10"))
mins$binsize <- paste(mins$widthx,"X",mins$widthx,sep="")

@@

<<FlightTradScatterplot, echo=F,include=F,eval=F, cache=TRUE>>=
source("RectangularBinningFunctions.R")
load(file="DepartDelaysMinutes.RData")
# Code takes a long long time to compile so ran once, saved to high res png, then set eval=FALSE
ggplot(aes(x=Scheduled/60, y=Actual/60), data=DepartDelaysMinutes) +
  geom_point(alpha=.05) + 
  theme_bw() + theme(aspect.ratio=1,
                     panel.grid= element_blank(),
                     legend.text=element_text(size=12) )  +
  scale_x_continuous("Scheduled Departure Time", limits=c(0,24), breaks=seq(0,24,by=6)) +
  scale_y_continuous("Actual Departure Time", limits=c(0,24), breaks=seq(0,24,by=6)) +
  ggtitle("Alpha-Blended Scatterplot of Airline Departure Times") 
ggsave(filename = "./figure/AlphaScatterAirline2014.png", width= 6, height=6, dpi=900)
@

<<FlightMinBinScat, echo=F,include=F, cache=TRUE>>=
source("RectangularBinningFunctions.R")
load(file="DepartDelaysMinutes.RData")
# Create minimal binning
departure1minute <- RectBin2d(xs=DepartDelaysMinutes$Scheduled,ys=DepartDelaysMinutes$Actual, 
                              originx = -0.5, originy = -0.5, 
                              widthx = 1, widthy = 1, type="standard")
# down to 3.261% of original number of rows
# nrow(departure1minute)/nrow(DepartDelaysMinutes)
ggplot(aes(x=binxs/60, y=binys/60), data=departure1minute) +
  geom_tile( aes(fill=binfreq) ) + 
  scale_x_continuous("Scheduled Departure Time", limits=c(0,24), breaks=seq(0,24,by=6)) +
  scale_y_continuous("Actual Departure Time", limits=c(0,24), breaks=seq(0,24,by=6)) +
  ggtitle("Minimally Binned Scatterplot of Airline Departure Times")+ 
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43", guide="legend", trans="log", breaks=c(1, 10, 100, 1000)) +
  theme_bw() + theme(aspect.ratio=1,
                   panel.grid= element_blank(),
                   legend.text=element_text(size=12) ) + 
  geom_abline(intercept=0,slope=1, color="gray70", linetype=3)
ggsave(filename = "./figure/MinimalBinnedAirline2014.png", width= 6.5, height=6, dpi=900)
@

>>>>

# connect <- dbConnect(dbDriver("MySQL"), host="mysql2.stat.iastate.edu", 
#                      port=3306, user="dbaccess", dbname="baseball")
# pitch <- new("dataDB", co=connect, table="Pitching")
# 
# d1 <- dbData(pitch, vars=c( "G", "SO")) # Games, Strike Outs
# 
# xtd.data <- function(x) {
# # replicate each line as often as given by the last entry
#   k <- dim(x)[2]
# 
#   res <- data.frame(sapply(1:(k-1), function(var) {
#   		return(rep(x[,var], x[,k]))
# 		}
# 	))
# 	names(res) <- names(x)[1:(k-1)]
# 	return(res)
# }
# d1.ext <- xtd.data(d1)
# dbDisconnect(connect)
# pitchdat <- d1.ext
# save(d1, d1.ext,pitchdat, file="./data/pitching.Rdata")
load("./data/pitching.Rdata")
@

@@

<<OriginOffsetSimulations, echo=F,include=T,eval=T,fig.width=8, fig.height=4, out.width='.9\\linewidth', fig.pos='h',fig.align='center',tidy=F, cache=TRUE, fig.show='hold',fig.cap="Net spatial loss for coarse simulated data at a 2X2 unit resolution using various sized square bins over the range of possible origin offsets. The vertical gray line in each facet indicates the origin offset of (1,1).">>= 
##  Check out the spatial loss for the data that is recorded to resolution = alpha at various origin offsets
# standard and random binning with bin sizes that are scalar multiples of the data resolution
# save spatial loss information
# spatlossdatRound <- data.frame(NULL)
# index=0
# distindex = 0
# distrib <- c("Uniform", "Normal", "Exponential")
# widths <- seq(alpha,5*alpha,by=alpha)
# for (dat in 1:3){
#   distindex <- distindex + 1
#   for (width in widths){
#     for (origin in seq(-width,0, by=width/60)){
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="standard")
#       binout[[2]]$bintype <- "standard"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="random")
#       binout[[2]]$bintype <- "random"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       print(c(distrib[distindex],width,origin))
#     }
#   }
# }
# write.csv(spatlossdatRound,"LossesByOriginBinwidth.csv", row.names=FALSE)
lossOrigin <- read.csv("./data/LossesByOriginBinwidth.csv",header=TRUE)
mins <- ddply(lossOrigin, .(widthx, bintype, distrib), summarise,
              minx = max(originx[which(totalSpatialLoss == min(totalSpatialLoss))]),
              minloss = min(totalSpatialLoss))
lossOrigin$binsize <- factor(paste(lossOrigin$widthx,"X",lossOrigin$widthx,sep=""),
                             levels=c("2X2","4X4","6X6","8X8","10X10"))
mins$binsize <- paste(mins$widthx,"X",mins$widthx,sep="")

@@

<<FlightTradScatterplot, echo=F,include=F,eval=F, cache=TRUE>>=
source("RectangularBinningFunctions.R")
load(file="./data/DepartDelaysMinutes.RData")
# Code takes a long long time to compile so ran once, saved to high res png, then set eval=FALSE
ggplot(aes(x=Scheduled/60, y=Actual/60), data=DepartDelaysMinutes) +
  geom_point(alpha=.05) + 
  theme_bw() + theme(aspect.ratio=1,
                     panel.grid= element_blank(),
                     legend.text=element_text(size=12) )  +
  scale_x_continuous("Scheduled Departure Time", limits=c(0,24), breaks=seq(0,24,by=6)) +
  scale_y_continuous("Actual Departure Time", limits=c(0,24), breaks=seq(0,24,by=6)) +
  ggtitle("Alpha-Blended Scatterplot of Airline Departure Times") 
ggsave(filename = "./figure/AlphaScatterAirline2014.png", width= 6, height=6, dpi=900)
@

<<FlightMinBinScat, echo=F,include=F, cache=TRUE>>=
source("RectangularBinningFunctions.R")
load(file="./data/DepartDelaysMinutes.RData")
# Create minimal binning
departure1minute <- RectBin2d(xs=DepartDelaysMinutes$Scheduled,ys=DepartDelaysMinutes$Actual, 
                              originx = -0.5, originy = -0.5, 
                              widthx = 1, widthy = 1, type="standard")
# down to 3.261% of original number of rows
# nrow(departure1minute)/nrow(DepartDelaysMinutes)
ggplot(aes(x=binxs/60, y=binys/60), data=departure1minute) +
  geom_tile( aes(fill=binfreq) ) + 
  scale_x_continuous("Scheduled Departure Time", limits=c(0,24), breaks=seq(0,24,by=6)) +
  scale_y_continuous("Actual Departure Time", limits=c(0,24), breaks=seq(0,24,by=6)) +
  ggtitle("Minimally Binned Scatterplot of Airline Departure Times")+ 
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43", guide="legend", trans="log", breaks=c(1, 10, 100, 1000)) +
  theme_bw() + theme(aspect.ratio=1,
                   panel.grid= element_blank(),
                   legend.text=element_text(size=12) ) + 
  geom_abline(intercept=0,slope=1, color="gray70", linetype=3)
ggsave(filename = "./figure/MinimalBinnedAirline2014.png", width= 6.5, height=6, dpi=900)
@
